Design and Implementation of Probabilistic Programming Language Anglican

### 2016-11-27 paper/paper.tex

\documentclass[preprint]{sigplanconf}

@@

An Anglican program, or \textit{query}, is compiled into a
Clojure function. When inference is performed with a provided
algorithm, this produces a sequence of samples. Anglican shares
a common syntax with Clojure; Clojure functions can be called
from Anglican code and vice versa. A simple program in Anglican
can look like the following code:
\begin{lstlisting}[style=default]
(defquery model data
  "chooses a distribution
   which describes the data"
  (let [;; Guessing a distribution
        dist (sample (categorical
                       [[normal 0.5]
                        [gamma 0.5]]))
        a (sample (gamma 1 1))
        b (sample (gamma 1 1))
        d (dist a b)]

    ;; Observing samples from the distribution
    (loop [data0 data] ;; Initializing data0
      (when (not-empty data0)
        (let [[x & data1] data0] 
          (observe d x)
          (recur data1))))

    ;; Returning the distribution and parameters
    [d a b]))
\end{lstlisting}
The query builds a model for the input $\texttt{data}$, 
a sequence of data points.
It defines a probability distribution on three variables, 
$\texttt{d} \in \{\texttt{normal},\texttt{gamma}\}$
for a distribution type, and $\texttt{a}$ and $\texttt{b}$ for
positive parameters for the type. Concretely,
using the $\texttt{sample}$ statements, the query
first defines a so called prior distribution on these variables.
Then, it adjusts this prior distribution based on data points in
$\texttt{data}$ using the \texttt{observe} statement.
Samples from this adapted distribution (also called posterior distribution)
can be obtained by running the query under one of Anglican's
inference algorithms.

>>>>

\documentclass[preprint]{sigplanconf}
\linespread{1.07}

@@

An Anglican program, or \textit{query}, is compiled into a
Clojure function. When inference is performed with a provided
algorithm, this produces a sequence of samples. Anglican shares
a common syntax with Clojure; Clojure functions can be called
from Anglican code and vice versa. A simple program in Anglican
can look like the following code:
\begin{minipage}{\linewidth}
\begin{lstlisting}[style=default]
(defquery model data
  "chooses a distribution
   which describes the data"
  (let [;;; Guess a distribution.
        dist (sample (categorical
                       [[normal 0.5]
                        [gamma 0.5]]))
        a (sample (gamma 1 1))
        b (sample (gamma 1 1))
        d (dist a b)]

    ;;; Observe samples from the distribution.
    (loop [observations data]
      (when (not-empty observations)
        ;; Retrieve the first observation as `o',
        ;; and store the rest of observations in 
        ;; 'observations*'.
        (let [[o & observations*] observations]
          ;; Observe 'o' from the guessed 
          ;; distribution 'd'.
          (observe d o))
        ;; Proceed to the next iteration with
        ;; the rest of observations.
        (recur observations*)))

    ;;; Return the distribution and parameters.
    [d a b]))
\end{lstlisting}
\end{minipage}
The query builds a model for the input $\texttt{data}$, 
a sequence of data points.
It defines a probability distribution on three variables, 
$\texttt{d} \in \{\texttt{normal},\texttt{gamma}\}$
for a distribution type, and $\texttt{a}$ and $\texttt{b}$ for
positive parameters for the type. Concretely,
using the $\texttt{sample}$ forms, the query
first defines a so called prior distribution on these three variables,
and then it adjusts this prior distribution based on observations in
$\texttt{data}$ using the \texttt{observe} form.
Samples from this adapted distribution (also called posterior distribution)
can be obtained by running the query under one of Anglican's
inference algorithms.
